### 2023-12-06 to_do.tex
to do list changes from ages ago

To Do:

Code:
* compressible Boussinesq equations class and examples
* general mixed solve for the compressible Euler equations
DONE* merge transport wrapper PR and update all examples and scripts
DONE?* fix vorticity function space
* check hydrostatic Euler is correct
* check SUPG is correct
* better provision of solver options?

Decisions:
* where to store scripts and how to archive them (GitHub repo + Zenodo)
* which simulations to run (see later section)
* will physics be part of this paper?

Website:
* dragon logo still apears in the top tab

Simulations:
* shallow water
 - lowest order and next-to-lowest order
 - cubed sphere and icosahedral sphere meshes
 - range of timestepping schemes: show convergence plot of W2
 - vector invariant vs advective form

Writing:
* describe advantages of building on top of Firedrake
* compressible Euler section
* compressible Boussinesq section
* Gusto design
* FML example
 - generate code
* future directions:
 - structure preserving discretisations / vorticity based discretisations
 - other equation sets
 - PDC
 - ML?
 - coupled models (mention Thetis and / or Icepack?)

Figures:
* cubed sphere and icosahedral sphere grids
* extruded mesh?
* DOFs?
* convergence plot of SW W2 (and W3?) for a range of timestepping schemes
* snapshots of W5 (after 15 and 50 days), W6 and Galewsky [suggest: W5 and W6 at both orders; Galewsky on both icosahedral and cubed spheres]
* vertical slice gravity wave final state
* vertical slice flow over orography
* dense bubble
* something in 3D